### 2024-08-05 docs/manuscript/manuscript.qmd
paper review update, still under review

---
title: ""
bibliography: "misc/ref.bib"
csl: "misc/vancouver.csl"
format: 
  docx: 
    reference-doc: "ref.docx"
---
**Title**: Physical Fitness is Related to Body Fat and Heart Rate Variability during Exercise in Community-Dwelling Older People: a cross-sectional study.

**Authors**: Diego Mabe-Castro^1,2^, Matías Castillo-Aguilar^1^, Matías Mabe-Castro^1,3^, Ruby Méndez Muñoz^1^, Carla Basualto-Alarcón^4,5^, Cristian Núñez-Espinosa^1,3,6^.

::: {custom-style="Afiliation"}
^1^ Centro Asistencial Docente e Investigación, Universidad de Magallanes, Punta Arenas, Chile

^2^ Departamento de Kinesiología, Universidad de Magallanes, Punta Arenas, Chile

^3^ Escuela de Medicina. Universidad de Magallanes, Punta Arenas, Chile

^4^ Laboratory of Cellular Physiology and Metabolism; Centro PUENTES de Investigación en Salud; Health Sciences Department, University of Aysén, Coyhaique, Chile

^5^ Anatomy and Legal Medicine Department, Faculty of Medicine, University of Chile, Santiago, Chile.

^6^ Interuniversity Center for Healthy Aging
:::

::: {custom-style="Afiliation"}
Corresponding Author:

Cristian Núñez-Espinosa

Avenida Bulnes 01855, Box 113-D, Punta Arenas, Región de Magallanes y la Antártica Chilena, Chile

Email address: Cristian.nunez\@umag.cl
:::

\newpage

## Abstract

**Background**. Age-related changes in body composition affect physical fitness in older adults. However, whether the autonomic response is associated with body fat percentage and its implication for physical fitness is not fully understood.

**Aim**. To understand the association between physical fitness, body composition, and heart rate variability in older people and its mediating factors.

**Methods**. A cross-sectional study with 81 older adults was conducted, assessing Short Physical Performance Battery (SPPB), Two-minute Step Test (TMST), body composition, and cardiac autonomic response. Correlation and mediation analyses were performed.

**Results**. Body fat percentage negatively correlated with physical fitness (SPPB: r = -0.273, p = 0.015; TMST: r = -0.279, p = 0.013) and sympathetic activity (SNS index: r = -0.252, p = 0.030), yet positively correlated with parasympathetic tone (RMSSD: r = 0.253, p = 0.029; SDNN: r = 0.269, p = 0.020). Physical fitness associated with sympathetic nervous system index (SPPB: r = 0.313, p = 0.006; TMST: r = 0.265, p = 0.022) and parasympathetic nervous system index (TMST: r = -0.344, p = 0.003). Muscle mass mediated body fat's impact on physical fitness, while physical fitness mediated body fat's impact on autonomic response.

**Conclusion**. Body composition and cardiac autonomic response to exercise are associated with physical fitness in older people, highlighting a possible protective effect of muscle mass against the decline in physical fitness associated with increased body fat.

\newpage

# Introduction

Aging is accompanied by a multitude of physiological changes that have the potential to significantly impact people's overall health and well-being [@pmid:30185958; @pmid:31154942]. Body composition and cardiovascular function are particularly relevant, as they are closely associated with aging [@pmid:36948782; @pmid:36443946; @pmid:30065258; @pmid:36918543; @pmid:32483625]. Recently, there has been a growing interest in comprehending the interplay between physical fitness, body composition, and cardiac autonomic response to exercise in older individuals, as these factors play pivotal roles in determining health outcomes within this age group.

<!---- Historical precedents ---->

Physical fitness is the ability to perform daily tasks with vigor, without undue fatigue, and with ample energy to enjoy leisure-time pursuits and meet unforeseen emergencies [@pmid:3920716]. Therefore, it is a crucial component of healthy aging. The term encompasses many aspects, including cardiorespiratory endurance, muscular endurance and strength, body composition, and flexibility [@pmid:3920711]. Numerous studies have demonstrated the positive impact of physical fitness on overall health and longevity, emphasizing its role in reducing the risk of chronic diseases and improving quality of life and functional independence in older adults [@pmid:9177581; @pmid:29293447; @pmid:30768197; @pmid:36332759]

Moreover, the alteration of body composition, specifically the proportion of body fat, is accompanied by significant changes with age, resulting in a tendency towards an increase in adiposity and a decrease in lean muscle mass [@pmid:35968789; @pmid:33673927]. Excessive body fat accumulation, particularly visceral adiposity, has been associated with a higher risk of cardiovascular disease, metabolic disorders, and functional limitations in older individuals [@pmid:31301983; @pmid:33882682].

The autonomic nervous system (ANS) plays a critical role in regulating cardiovascular function, with sympathetic and parasympathetic branches exerting opposing effects on heart rate and vascular tone. Furthermore, during stressful situations, such as physical exercise, the ANS ensures a sufficient cardiac response to higher metabolic demands [@pmid:16627049]. Heart rate variability (HRV) is a non-invasive indicator of ANS activity and cardiovascular health [@pmid:37641564; @pmid:29033120]. A reduction in resting HRV has been linked to various adverse outcomes, including cardiovascular events and morbidity, while insufficient HRV reduction during exercise may lead to impaired physical capacity [@pmid:29033120; @pmid:33390146; @pmid:36054204].

However, HRV is not only influenced by physical and environmental stressors. Psychological factors such as anxiety and depression have also been linked to changes in ANS activity [@doi:10.1016/j.jad.2018.04.071; @pmid:35340102]. Older individuals may be more susceptible to these psychological factors, making it crucial to measure and control for their effects [@pmid:38007499; @pmid:25980510].

<!---- Conceptualization of the issue ---->

Despite extensive research into physical fitness, body composition, and cardiovascular health in older adults, gaps persist in understanding their interrelationships and the factors mediating them. One area of interest is the relationship between physical fitness, body fat percentage, and cardiac autonomic response to exercise in older individuals. This study explores their collective influence on physical fitness in the aging population.

The conceptualization of this investigation is based on recognizing the interplay between physical fitness, body composition, and autonomic cardiovascular regulation in aging. We aim to answer the following research question: How does physical fitness relate to body fat percentage and cardiac autonomic responses to exercise in older people? By elucidating these correlations, we hope to gain valuable insights into the relationships that underlie age-related modifications in physical fitness and identify potential interventions to enhance health outcomes in older adults. Even more, we seek to understand the mediating effect of the variables, including physical aspects and frequent psychological traits, such as geriatric depression and anxiety.

<!---- Hypothesis and main objectives ---->

We hypothesized that physical fitness would be inversely related to body fat percentage and associated with cardiac autonomic response during exercise in older individuals. We believe an optimal cardiac autonomic response, characterized by a shift towards sympathetic activity predominance, will be essential for this population to achieve greater performance during physical fitness testing.

# Methods

## Aims and Study design

A prospective observational, cross-sectional study was conducted in one data collection session to determine the associations between physical fitness, body composition, and cardiac autonomic response to moderate-intensity exercise.

## Setting

This study was conducted at the Centro Asistencial Docente e Investigación (Assistance, Teaching, and Research Center), which belongs to the University of Magallanes (CADI-UMAG) in Punta Arenas, Chile. All the assessments were made between 09:00 and 11:00 a.m. The privacy and comfort of the subjects were ensured; the room temperature was set at 20°C, and white artificial lighting was used.

## Participants

81 community-dwelling older adults were recruited and selected by non-probabilistic sampling from the Region of Magallanes and Chilean Antarctica, Chile. If they were 60 years or older, residents of the Region of Magallanes and Chilean Antarctica, Chile, they were included and understood the study aims and instructions. However, they were excluded in the case of diagnosis of congenital heart disease, consumption of beta-blocker drugs, taking stimulant substances within 24 h before assessment session, motor or cognitive disability, inability to understand instructions or written content, or presence of pain during cardiac or physical assessments.

All participants gave their permission and provided informed written consent before participation. The Ethics Committee of the University of Magallanes (N°10/CEC-UMAG/2023) approved this study, following the regulations established by the Declaration of Helsinki on ethical principles in human beings.
The flow diagram of the included participants can be seen in Figure 1.

## Procedures

During recruitment, the participants were instructed to avoid the use of psychoactive substances for 24 hours before the assessment and to sleep for at least 7 hours the previous night. Upon arrival, participants were informed about the study aims and risks associated with their participation, and sociodemographic and medical information was collected during the initial interview. Then, Body composition parameters were measured using bioimpedance analysis, and psychological questionnaires for anxiety and depression screening were applied by a supervised psychology undergraduate student.

Cardiac autonomic response to physical exercise was measured as previously validated for this age group [@ijerph20054456]. The protocol is briefly described below:

HRV was recorded through a non-invasive chest band immediately before, during, and after executing the two-minute step test (TMST). Vital signs, including blood pressure, were monitored throughout the test, and participants’ well-being was visually checked to ensure they were comfortable and prepared. For resting HRV measurements (before and after the application of the TMST), the volunteers remained seated in a chair, with feet and back supported, ensuring avoidance of talking during the recordings. R-R intervals were recorded continuously during the last 10 min of rest and were analyzed for 5 min on each occasion. The breathing rate was spontaneous. As part of the protocol, it was ensured that the participant had a blood pressure of less than 140/90 mmHg to start the HRV measurements.

Finally, after 15 minutes of resting from the TMST, the Short Physical Performance Battery (SPPB) was administered to the participants. Before the session ended, the subjects ' blood pressure, heart rate, and general appearance were verified.

Physical and physiological assessments, as well as the initial interview, were made by professional physiotherapists.

## Instruments

### Short physical performance battery

SPPB consists of a physical test used to measure three components of physical fitness, described below [@Guralnik1994]:

Balance: to achieve the maximal score (4 points), the subject should be able to stay balanced for at least 10 seconds in a side-by-side stand (feet together), semi-tandem stand, and tandem stand.

Usual gait speed: The subjects are asked to walk normally at a 4-meter distance. They are given two chances, and the best of both is registered. To achieve the maximal score (4 points), they should obtain 4.82 sec. or less.

Lower body fitness: The subjects are asked to perform five chair stands without using their arms, and time is registered upon completion. A time lower or equal to 11.19 sec. gives the maximal score (4 points).

To obtain the final score, the sum of the three components is calculated, with a maximal total score of 12 points.

### Body composition

Body mass (kg) and total body fat (%) were assessed by bioimpedance using the Tanita BC-558 Ironman Segmental Body Composition Monitor (Tanita Ironman, Arlington Heights, IL 60005, USA), with a concordance of 89.3% compared to the Dual X-ray Absorption test using standard measurement protocols [@Mialich2011].

### Cardiac autonomic activity

The cardiac autonomic activity was assessed using R-R interval recordings (HRV) obtained through the Polar Team2 system (Polar®) application. Artifacts and ectopic heartbeats were excluded, not exceeding 3% of the recorded data [@malik1996heart]. Time-domain parameters considered for analysis included the square root of the mean squared differences of successive R-R intervals (RMSSD, expressed in ms) as an index of parasympathetic activity [@pmid:20033207], and the standard deviation of RR intervals (SDNN), reflecting total variability encompassing both sympathetic and parasympathetic contributions to cardiac autonomic function [@pmid:9401419; @buchheit2006cardiac]. The Stress Index (SI) and Parasympathetic and Sympathetic Nervous System Index (PNS and SNS) were computed. The PNS Index, indicative of total vagal stimulation, was derived from mean R-R intervals, RMSSD, and Poincaré Plot Index SD1 in normalized units (linked to RMSSD), representing deviations from normal population averages [@pmid:9401419; @GSID:XkAQwTGLgNsJ]. The SNS Index, reflecting total sympathetic stimulation, was derived from mean R-R intervals, Baevsky’s Stress Index (positively related to cardiovascular system stress and cardiac sympathetic activity), and the Poincaré Plot Index SD2 in normalized units (related to SDNN) with interpretation similar to the PNS Index [@pmid:9401419; @GSID:XkAQwTGLgNsJ]. The SI indicates the ANS control system’s workload [@doi:10.1159/000513781], normalized by the square root of Baevsky’s SI [@2009MethodicalRU]. All analyses were conducted to compute HRV-related indices using Kubios HRV® software (Kuopio, Finland).

### Two-Minutes Step Test

The TMST is a sub-test from the Senior Fitness Test, demanding a low to moderate intensity [@Rikli1999; @ijerph20054456]. It consists of a two-minute assessment designed to evaluate cardiorespiratory fitness. Participants are instructed to raise their right knee to a marked point on a wall as often as possible within the given time frame, ensuring that each raise reaches at least a 70° angle at the thigh-femoral joint. The number of valid steps was recorded for each subject.

### Geriatric Depression Scale

The 30-question Geriatric Depression Scale (GDS-30) was employed to assess the subject’s depressive symptoms [@pmid:27633329]. It consists of a dichotomous questionnaire, where participants are asked about their past-week feelings about depressive symptoms (for instance, “Do you feel that your life is empty?”), with higher scores (i.e., “yes” responses) representing more depressive symptoms. It was first developed by Yesavage and colleagues [@pmid:7183759] and is currently widely used [@pmid:34922595]. The Spanish version employed in this study has been previously validated, with a Cronbach alpha coefficient of 0.82 [@pmid:11921157]. The authors have permission to use this instrument from the copyright holders.

### Beck Anxiety Inventory

The Beck Anxiety Inventory was employed to assess anxiety symptoms. It consists of a 21-item questionnaire based on usual anxiety symptoms and a 4-option Likert scale from 0 (“Not at all”) to 3 (“Severely”), meaning the severity reported by the subject in each one. It was originally developed by Beck and colleagues [@pmid:3204199]. This study uses the Spanish version, demonstrating a high internal consistency in older people (α = 0.94) [@RODRIGUEZREYNALDO2001353]. The authors have permission to use this instrument from the copyright holders.

## Statistical analysis

We used mean and standard deviation (SD) to describe continuous variables and absolute and relative frequencies to describe discrete variables. We used Pearson’s product-moment correlation (r) to assess the relationship between variables. To assess differences between groups, we used standardized mean difference (SMD) and 95% confidence intervals (CI!95%!).

As a way of controlling for the influence of psycho-physiological variables on the cardiac autonomic response of the subjects to exercise, we assess the average causal mediation effect (ACME), the average direct effect (ADE) of the main effects after taking into account the effect of moderator variables into the observed relationships and the proportion of the effect that its mediated by these variables [@kosuke2010identification]. TTo assess the significance of moderating variables, we used nonoverlapping CI95%, estimated based on nonparametric bootstrapping using Monte Carlo resampling, and bias-corrected. We accelerated CI95%, using the *mediation* R package to this end [@tingley2014mediation].

We defined a type I error rate of 5% (α = 0.05) as our threshold for null hypothesis significance testing and nonoverlapping CI95% over the null effect for sex differences and mediation analyses. To estimate the confidence in our conclusions and, consequently, the statistical power of our analyses, we calculated the probability of correctly rejecting the null hypothesis for the correlation tests post hoc, considering a low (i.e., r = 0.3) to moderate effect size (r = 0.5)

All analyses were computed using the R programming language for statistical computing [@rlanguage].

# Results

## Sample characteristics

The sample consisted of 81 adults (71.1 ± 6.2 years old, BMI: 31 ± 6.2 kg/m^2^) and 67 (82.7%) females. Table 1 shows the characteristics of the sample.

```{r prep-env}
#| include: false
#| message: false
#| warning: false
library(flextable)
library(data.table)
library(ggplot2)
load("../../data/mindfulness.RData")
dat <- mindfulness[id_event_name == "Basal", -2L]
```

```{r}
#| eval: false
#| echo: false
#| message: false
#| warning: false
readRDS(file = "tables/tbl-descriptives.RDS") |> 
  font(part = "all", fontname = "Times")
```

## Body composition and autonomic response

We observed that body fat percentage was linked to a lower SNS (r = -0.252, p = 0.030) and Stress index (r = -0.258, p = 0.027) response during the TSMT. In addition, body fat percentage was positively correlated with time domain HRV during exercise, specifically RMSSD (r = 0.253, p = 0.029) and SDNN (r = 0.269, p = 0.020) (See Figure 2). No other body composition variables correlate with the cardiac autonomic response to exercise.

```{r}
#| echo: false
#| warning: false
#| error: false
#| fig-width: 8
#| fig-height: 4
#| fig-dpi: 300
dat_long <- melt.data.table(
  data = dat,
  id.vars = c("id_record", "cc_grasa_total_porcentaje"), 
  measure.vars = c("hrv_sns_peri", "hrv_stress_peri", "hrv_rmssd_peri", "hrv_sdnn_peri", "hrv_pns_peri")
)

dat_long[, variable := fcase(
  grepl("sns", variable), "SNS index",
  grepl("pns", variable), "PNS index",
  grepl("stress", variable), "Stress index",
  grepl("sdnn", variable), "SDNN",
  grepl("rmssd", variable), "RMSSD"
)]

dat_long[, variable := factor(variable, levels = c("RMSSD", "SDNN", "PNS index", "SNS index", "Stress index"))]

ggplot(dat_long, aes(value, cc_grasa_total_porcentaje)) +
  facet_grid(~ variable, scales = "free_x") +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_text(mapping = aes(label = text), data = data.frame(
    variable = as.factor(c("RMSSD", "SDNN", "PNS index", "SNS index", "Stress index")),
    cc_grasa_total_porcentaje = c(61, 61, 61, 61, 61),
    value = c(13, 20, -2, 8, 45),
    text = c("p = 0.029", "p = 0.020", "p = 0.262", "p = 0.030", "p = 0.026")
  ), size = 3.5, fontface = "italic") +
  labs(y = "Body fat (%)", x = "Values") +
  tidybayes::theme_ggdist()
```

**Figure 1**. Bivariate dispersion plots between body fat and HRV related measures. Significance values for the Pearson's product-moment correlation test is shown.

## Body composition and physical fitness

In addition, greater SPPB score was observed in leaner (r = -0.273, p = 0.015) and individuals with lower BMI values (r = -0.225, p = 0.047). Similar findings were observed between total steps from the TMST and body fat percentage (r = -0.279, p = 0.013) and bodyweight (r = -0.232, p = 0.041). Not enough evidence was observed to suggest a correlation between total muscle mass with either total SPPB score (r = -0.063, p = 0.584) or total number of steps in TMST (r = -0.070, p = 0.541).

```{r}
#| echo: false
#| warning: false
#| error: false
#| fig-width: 8
#| fig-height: 4
#| fig-dpi: 300
dat_long <- melt.data.table(
  data = dat,
  id.vars = c("id_record", "cc_grasa_total_porcentaje", "cc_masa_muscular_total", "sd_sexo"), 
  measure.vars = c("tm_pasos", "sppb_total")
)

dat_long[, variable := fcase(
  grepl("tm_pasos", variable), "TMST steps",
  grepl("sppb_total", variable), "SPPB total score"
)]

dat_long[, variable := factor(variable, levels = c("TMST steps", "SPPB total score"))]

fig2a <- ggplot(dat_long, aes(value, cc_grasa_total_porcentaje)) +
  facet_grid(~ variable, scales = "free_x") +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_text(mapping = aes(label = text), data = data.frame(
    variable = as.factor(c("TMST steps", "SPPB total score")),
    cc_grasa_total_porcentaje = c(60, 60),
    value = c(42, 4),
    text = c("p = 0.013\nr = -0.279", "p = 0.015\nr = -0.273")
  ), size = 3.5, fontface = "italic") +
  labs(y = "Body fat (%)", x = NULL) +
  scale_y_continuous(expand = c(.1,.1)) +
  tidybayes::theme_ggdist()

fig2b <- ggplot(dat_long, aes(value, cc_masa_muscular_total)) +
  facet_grid(~ variable, scales = "free_x") +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_text(mapping = aes(label = text), data = data.frame(
    variable = as.factor(c("TMST steps", "SPPB total score")),
    cc_masa_muscular_total = c(65.5, 65.5),
    value = c(42, 4),
    text = c("p = 0.541\nr = -0.070", "p = 0.584\nr = -0.063")
  ), size = 3.5, fontface = "italic") +
  labs(y = "Muscle mass (kg)", x = NULL) +
  scale_y_continuous(expand = c(.1,.1)) +
  tidybayes::theme_ggdist()

ggpubr::ggarrange(
  fig2a, fig2b, nrow = 1
)
```
**Figure 2**. Bivariate dispersion plots between physical fitness and body composition related measures. Significance values for the Pearson's product-moment correlation test is shown.

## Physical fitness and cardiac autonomic response to exercise

Physical fitness, reflected through SPPB total score and total steps in TMST, was positively correlated with a greater sympathetic activity during exercise, which was reflected by SNS index (SPPB score, r = 0.313, p = 0.006; TMST steps, r = 0.265, p = 0.022), mean HR (SPPB score, r = 0.355, p = 0.002; TMST steps, r = 0.338, p = 0.003) and its inverse counterpart mean R-R (SPPB score, r = -0.35, p = 0.002; TMST steps, r = -0.311, p = 0.007). Total number of steps from TMST was also inversely correlated with parasympathetic indices during exercise like RMSSD (r = -0.285, p = 0.014) and PNS index (r = -0.344, p = 0.003).

However, this appears to be accompanied by an proportional decrease in parasympathetic drive during exercise with increasing levels of physical fitness displayed in the SPPB, which was denoted by the inverse correlation of the later with PNS index during (r = -0.375, p = 0.001) and after exercise (r = -0.229, p = 0.049), in addition to RMSSD during exercise (r = -0.294, p = 0.010)

```{r}
#| echo: false
#| warning: false
#| error: false
#| fig-width: 8
#| fig-height: 8
#| fig-dpi: 300
dat_long <- melt.data.table(
  data = dat,
  id.vars = c("id_record", "tm_pasos", "sppb_total", "sd_sexo"), 
  measure.vars = c("hrv_sns_peri", "hrv_stress_peri", "hrv_rmssd_peri", "hrv_sdnn_peri", "hrv_pns_peri")
)

dat_long[, variable := fcase(
  grepl("sns", variable), "SNS index",
  grepl("pns", variable), "PNS index",
  grepl("stress", variable), "Stress index",
  grepl("sdnn", variable), "SDNN",
  grepl("rmssd", variable), "RMSSD"
)]

dat_long[, variable := factor(variable, levels = c("RMSSD", "SDNN", "PNS index", "SNS index", "Stress index"))]

fig3a <- ggplot(dat_long, aes(value, tm_pasos)) +
  facet_grid(~ variable, scales = "free_x") +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_text(mapping = aes(label = text), data = data.frame(
    variable = as.factor(c("RMSSD", "SDNN", "PNS index", "SNS index", "Stress index")),
    tm_pasos = c(125, 125, 125, 125, 125),
    value = c(13, 20, -2, 8, 45),
    text = c("p = 0.014\nr = -0.285", "p = 0.141\nr = -0.173", "p = 0.002\nr = -0.344", "p = 0.022\nr = 0.265", "p = 0.261\nr = 0.132")
  ), size = 3.5, fontface = "italic") +
  labs(y = "TMST steps", x = NULL) +
  scale_y_continuous(expand = c(.1,.1)) +
  tidybayes::theme_ggdist() +
  theme(axis.line.x = element_blank(), axis.ticks.x = element_blank(), 
        axis.text.x = element_blank())

fig3b <- ggplot(dat_long, aes(value, sppb_total)) +
  facet_grid(~ variable, scales = "free_x") +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_text(mapping = aes(label = text), data = data.frame(
    variable = as.factor(c("RMSSD", "SDNN", "PNS index", "SNS index", "Stress index")),
    sppb_total = c(14, 14, 14, 14, 14),
    value = c(13, 20, -2, 8, 45),
    text = c("p = 0.010\nr = -0.294", "p = 0.419\nr = -0.095", "p < 0.001\nr = -0.375", "p = 0.006\nr = 0.313", "p = 0.095\nr = 0.194")
  ), size = 3.5, fontface = "italic") +
  labs(y = "SPPB score", x = "Values") +
  scale_y_continuous(expand = c(.1,.1)) +
  tidybayes::theme_ggdist()

ggpubr::ggarrange(
  fig3a, fig3b, ncol = 1
)
```
**Figure 3**. Bivariate dispersion plots between physical fitness and HRV related measures. Significance values for the Pearson's product-moment correlation test is shown.

## Mediation analysis

Mediation analyses suggest many potential influential effects for the relationships for body composition and parasympathetic indices. In this context, the average direct effect (ADE) of body fat percentage on RMSSD during exercise (ADE = 0.088, CI~95%~[-0.005, 0.200], p = 0.077) was accentuated when considering the influence of SPPB score (ACME = 0.0289, CI~95%~[0.003, 0.110], p = 0.057; Total effect = 0.117, CI~95%~[0.028, 0.240], p = 0.012). Similar findings were observed when considering the mediation effect of TMST steps into account (ACME = 0.0283, CI~95%~[0, 0.07], p = 0.079; Total effect = 0.117, CI~95%~[0.028, 0.240], p = 0.012). No other parasympathetic indicators were influenced or mediated by fitness or psychological related measures.

Additionally, and in the case on sympathetic indicators, the effect of body fat on SNS index during exercise (ADE = -0.050, CI~95%~[-0.107, 0.01], p = 0.077) was influenced by SPPB score in similar way as with RMSSD, enhancing the original main observed effect (ACME = -0.016, CI~95%~[-0.049, 0], p = 0.041; Total effect = -0.066, CI~95%~[-0.125, -0.010], p = 0.016). Similar effects were observed with TMST steps in this regard (ACME = -0.015, CI~95%~[-0.042, 0], p = 0.050; Total effect = -0.066, CI~95%~[-0.125, -0.010], p = 0.016). No other sympathetic indices were notoriously modified in the presence of either SPPB score, TMST steps or psychological variables.

When assessing potential mediators for the effect of body composition on fitness related measures, we found that the effect of body fat percentage on SPPB score (ADE = -0.075, CI~95%~[-0.123, -0.020], p = 0.007) is partially diminished when considering the influence of total muscle mass (ACME = 0.011, CI~95%~[0, 0.040], p = 0.107; Total effect = -0.064, CI~95%~[-0.111, -0.010], p = 0.014). 
Similar mediation effects of total muscle mass were observed (ACME = 0.118, CI~95%~[0.005, 0.350], p = 0.075) when assessing the effect of body fat percentage on TMST steps (ADE = -0.773, CI~95%~[-1.193, -0.300], p = 0.001; Total effect = -0.655, CI~95%~[-1.084, -0.210], p = 0.004). No other mediating effects were observed for any of the psychological measures for any of the relationships between body composition and physical fitness measures.

![](figures/fig_esquema.jpg)
**Figure 4**. Diagram of main results observed form the data. Blue arrows and boxes represent the correlation between variables, while orange arrows and boxes represent the factors mediating the above correlations. "+" means that the mediating factor enhances the main observed effect, contrary to "-", reflecting a diminished main effect when considering the mediating factor.

# Discussion

This study aimed to investigate the relationship between physical fitness, body fat percentage, and cardiac autonomic response to exercise in older people. Through correlational and mediating analyses, we aimed to elucidate the interplay between these factors and their implications for cardiac autonomic function during physical efforts.

It was hypothesized that a higher body fat percentage would be inversely correlated to physical fitness. Our results aligns with this hypothesis and with those of previous research regarding body composition and physical performance, with a proportional decrease in aging individuals [@pmid:35832415; @pmid:36256894].

However, the mediating analysis showed that total muscle mass diminished the main observed effect of body fat on physical fitness, with more muscled individuals displaying a better physical performance despite the body fat percentage. This finding suggests that older adults could be benefited of activities that promotes their muscle mass conservation, even when they have high body fat percentages [@pmid:27663845].  Nevertheless, caution is advised in the interpretation of this result, as the close physiological association among increasing body fat and decreasing muscle mass during aging have been well-studied [@pmid:33335909]

Regarding the link between physical fitness and cardiac autonomic response to exercise, our results evidenced that individuals with a high sympathetic drive while exercising, had greater physical fitness. Previous research had demonstrated the utility of HRV and exercise-induced responses in those metrics as a potential marker of cardiorespiratory fitness in these individuals [@pmid:36054204], supporting the current hypothesis that a greater sympathetic drive facilitates physical performance during moderate intensity activities. Even more, it has been previously demonstrated that physically active individuals have an enhanced autonomic response to exercise, playing a pivotal role and strengthening this hypothesis [@pmid:28721030; @pmid:24799513].

Furthermore, the increase in body fat percentage was related to a reduced sympathetic activity and a higher parasympathetic activity during TMST, as hypothesized (i.e., SNS index, SI, SDNN and RMSSD). Throughout the course of physical efforts, ANS is expected to shift into a sympathetic predominance, ensuring a sufficient response to higher metabolic demands [@pmid:33246163]. However, aged individuals may present an impaired ANS response to physical exercise [@ijerph20054456]. In that sense, our results support that body composition plays a role in autonomic response. Moreover, they suggest that the negative effects of body fat on physical fitness in older adults, may be partially attributable to less efficient autonomic inputs to the cardiovascular system [@pmid:36875496].

Surprisingly, mediating analyses found that higher physical fitness (SPPB and TMST scores) intensifies the negative impact of body fat on autonomic control. This unexpected finding hints at a complex interplay between physical fitness, adiposity, and cardiac autonomic regulation. Our results indicate that in individuals with greater functional capacity, the influence of body fat on HRV is amplified rather than mitigated. This observation challenges traditional conceptions about the relationship between physical fitness and cardiac autonomic regulation, and raises questions about the underlying mechanisms involved.

As far as we know, there are not preceding studies addressing these findings. We posit that there exists further factors influencing the mediation effect. For instance, body distribution of the adipose tissue was not explored, specially when it is known that visceral fat has an important impact on cardiovascular health and its autonomic regulation [@pmid:31553779; @pmid:32158768]. Participants with higher fitness may have a different fat distribution compared with lower performance individuals, which may induce different effects on autonomic drive to the heart [@pmid:19752873]. Furthermore, other physical physiological, demographic or psychological variables may be playing an unexplored role. However, these hypothesis must be examined under scientific standards. Therefore, exhaustive research is needed to address this intriguing interplay.

Our results provide evidence to support the interplay between physical fitness, body fat, and cardiac autonomic response to exercise in older individuals. Physical fitness is negatively affected by higher body fat percentage, while also being associated with impaired cardiac autonomic response to exercise, characterized by decreased sympathetic drive and increased parasympathetic tone. Additionally, this altered autonomic response is related to lower physical performance in older individuals, both variables being important predictors of health outcomes in this population [@pmid:30627571; @pmid:30104034]. Overall, our findings underscore the triangular and bidirectional relationship among these variables.

However, our results did not evidence a mediating effect of either GDS-30 or BAI scores, challenging our initial hypothesis. These finding suggests a nuanced interplay between psychological variables and physical/physiological measures in older people, underscoring the need for further research to elucidate underlying mechanisms and variables beyond depression and anxiety. While the lack of evidence suggesting any mediating effects does not diminish the significance of psychological factors in aging, it highlights the complexity of their influence.

This study was strengthened by the fact that both SPPB and TMST measures consistently aligned in their interaction with other variables, which supports a global understanding of physical fitness in older people, rather than specific aspects, and by the exploration of mediating factors in the primary results. Nonetheless, this cross-sectional study, it is not without limitations. First, our study design prevents us from making causal inferences about the direction of the relationships, highlighting the relevance of experimental and longitudinal research exploring causal relations between our variables and potential underlying mechanisms. Second, a relatively small sample was included, limiting generalizability and statistical power of the analysis. Furthermore, the sample consisted mainly of women (82.7%), limiting our ability to make sex comparisons and decreasing the representativeness of the sample. Further research could investigate the impact of cardiac autonomic response on performance during the same task, in this age group.

# Conclusion

The results highlighted a structural bidirectional relationship among variables. Body fat percentage was inversely correlated to both physical fitness and cardiac autonomic activity during exercise, with sympathetic drive associated with physical fitness in older people. Furthermore, physical fitness may mediate the effect of body fat on cardiac autonomic activity during exercise, and total muscle mass may mitigate the negative effect of increased adiposity on physical fitness, highlighting its pivotal role in older people's health. However, further research is needed to elucidate mediating effects of psychological variables in this field.

# Author Contributions

All authors listed have made a substantial, direct and intellectual contribution to the work and approved it for publication.

# Funding

[...].

# Institutional Review Board Statement

[...].

# Informed Consent Statement

[...]

# Data availability statement

The raw data supporting the conclusions of this article will be made available by the authors without undue reservation.

# Conflicts of interests

The authors declare that the research was conducted without any commercial or financial relationships construed as a potential conflict of interest.

# Referencias

<div id="refs"></div>